---
title: "Book Notes"
subtitle: "Statistical Learning with R"
author: "Jon Geiger"
date: \today
output: pdf_document
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tentative Schedule

Winter 2022: 

- Chapter 1: Introduction

- Chapter 2: Statistical Learning

- Chapter 4: Classification

- Chapter 5: Resampling Methods

- Chapter 6: Linear Model Selection \& Regularization

- Chapter 8: Tree-Based Methods

Spring 2022: 

- Chapter 7: Moving Beyond Linearity

- Chapter 9: Support Vector Machines

- Chapter 12: Unsupervised Learning

# Organization

Chapters get single-# headers, and sections get double-## headers. 

Sub-sections (e.g. Chapter 2.1.3) are separated with horizontal lines (created with three asterisks):  

***


\clearpage

# Chapter 1: Introduction
```{r child = 'ch01/ch1_notes.Rmd'}
```

\clearpage

# Chapter 2: Statistical Learning
```{r child = 'ch02/ch2_notes.Rmd'}
```

\clearpage

# Chapter 4: Classification
```{r child = 'ch04/ch4_notes.Rmd'}
```

\clearpage

# Chapter 5: Resampling Methods
```{r child = 'ch05/ch5_notes.Rmd'}
```

\clearpage

# Chapter 6: Linear Model Selection and Regularization
```{r child = 'ch06/ch6_notes.Rmd'}
```

\clearpage

# Chapter 8: Tree-Based Methods
```{r child = 'ch08/ch8_notes.Rmd'}
```