---
title: "Book Notes"
subtitle: "Statistical Learning with R"
author: "Jon Geiger"
date: \today
output: pdf_document
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tentative Schedule

**Week 1**: Chapter 1

**Week 2**: Chapter 2.1-2.2

**Week 3**: Chapter 4.1-4.3 (4.4 optional). Lab 4.7.1-2 (4.7.3-5 optional).

**Week 4**: Chapter 4.5-4.6. Lab 4.7.7.

**Week 5**: Chapter 5.1-5.2. Lab 5.3.1-4.

**Week 6**: Chapter 6.2-6.3 (6.4 optional). Lab 6.5.1-3.

**Week 7**: Chapter 8.1-8.2. Lab 8.3.1-5

**Week 8**: Chapter 12.1-12.2. Lab 12.5.1.

**Week 9**: Chapter 12.4. Lab 12.5.3-4

**Week 10**: Grace Week

# Organization

Chapters get single-# headers, and sections get double-## headers. 

Sub-sections (e.g. Chapter 2.1.3) are separated with horizontal lines (created with three asterisks):  

***


\clearpage

# Chapter 1: Introduction
```{r child = 'ch1/ch1_notes.Rmd'}
```

\clearpage

# Chapter 2: Statistical Learning
```{r child = 'ch2/ch2_notes.Rmd'}
```

\clearpage

# Chapter 4: Classification
```{r child = 'ch4/ch4_notes.Rmd'}
```

\clearpage

# Chapter 5: Resampling Methods
```{r child = 'ch5/ch5_notes.Rmd'}
```

\clearpage

# Chapter 6: Linear Model Selection and Regularization
```{r child = 'ch6/ch6_notes.Rmd'}
```